---
title: "R Tutorial Basic"
author: "Md Delwar Hossain"
date: "2022-11-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R Tutorial
### 1. R Syntax
```{r}
"Hello World"
5
6
7

5 + 5

print("Delwar")

for (x in 1:10) {
  print(x)
}
```
### 2. R Comments

```{r}
7 + 3 #This is a comment

"Delwar" # Another Comment
```

### 3. R Variables

```{r}
name <- 'Delwar'
age <- 30
name
age
```

### 4. Concatenate Elements

```{r}
paste(name, age)

text <- "I Love my Society"
paste("Ok ", text)

num1 <- 5
num2 <- 10

num1 + num2
```


### 5. Multiple Variable

```{r}
x <- y <- z <- 10
x
y
z
```

### 6. Legal Variable Names

```{r}
myvar = "Delwar"
my_var = "Delwar"
myVar = "Delwar"
MYVAR = "Delwar"
myvar2 = "Delwar"
.myvar = "Delwar"
```

### 7. Data Types

```{r}
x <- 20
class(x)

y <- "Salay"
class(y)

i <- 9 + 4i
class(i)

z <- TRUE
class(z)

t = 10L
class(t)

c = 5i
class(c)
```

### 8. Type Conversion

```{r}
#You can convert from one type to another type with the following functions
x <- 4L
y <- 5
class(x)
class(y)

a <- as.numeric(x)
b <- as.integer(y)

class(a)
class(b)
```

### 9. R Math

```{r}
10 + 5
10 - 4
```

### 10. R Built in math function

```{r}
max(5, 10, 15)
min(5, 10, 15)

sqrt(16)

abs(-10) # return positive number

ceiling(1.6)
ceiling(1.3)

floor(1.6)
floor(1.3)
```

### 11. R String

```{r}
str = 'I am Delwar'
str1 <- 'I am HP'
str
str1

str2 <- 'I am writing a letter,
I am doing my duties,
You can buy someting
Learn R programming'

str2

cat(str2)
```

### 12. String Length

```{r}
nchar(str)
nchar(str2)
```

### 13. Check a string

```{r}
grepl('am', str2)
grepl('not', str2)
grepl('buy', str2)
```

### 14. Escape character to avoid error

```{r}
str <- 'I am not working "hard", in my job'
str
cat(str)
```

### 15. R Boolean 

```{r}
10 > 9
10 == 0
10 < 9

a <- 10
b <- 20
a > b

if (a > b) {
  print(a)
} else{
  print(b)
  
}
```

### 16. R Oprators

```{r}
a = 10
b = 20

a + b
a - b
a * b
a / b
a %% b
a^b 
```

### 17. R Comparison Operators

```{r}
a == b
a != b
a > b
a < b
a >= b
a <= b
```

### 18. R Logical Operators

```{r}
a & b
a && b

x = 1:10
x
```

### 19. R if....else

```{r}
a <- 5
b<- 10

if (a > b) {
  print("A is Greater than B")
}else{
  print("B is Greater than A")
}
```

### 20. elseif 

```{r}
a <- 5
b<- 10

if (a > b) {
  print("A is Greater than B")
}else if(a == b){
  print("A is equal to B")
}else{
  print("None")
}
```

### 21. Nested elseif 

```{r}
x <- 41

if (x > 10) {
  print("Above ten")
  if (x > 20) {
    print("and also above 20!")
  } else {
    print("but not above 20.")
  }
} else {
  print("below 10.")
}

```

### 22. Use of AND/OR

```{r}
a <- 200
b <- 33
c <- 500

if (a > b & c > a) {
  print("Both conditions are true")
}

if(a>b | c>a){
    print("At least one of the conditions is true")
}


```

### 23. R While Loops

```{r}
i <- 1
while (i<6) {
  print(i)
  i <- i+1
}
```
```{r}
i <- 1
while (i<6) {
  print(i)
  i <- i+1
  if(i==4){
    break
  }
}

```
With the next statement, we can skip an iteration without terminating the loop:
```{r}
i <- 1
while (i<6) {
  i <- i+1
  if(i==3){
    next
  }
  print(i)
}
```
```{r}
 dice <-1

while (dice<=6) {
  if(dice<6){
    print("Not Allowed")
  }else{
    print("Allowed")
  }
  dice<- dice + 1
}
```

### 24. R For Loop

```{r}
for (x in 1:5) {
  print(x)
}
fruits <- list("apple","banana","cherry")
for (i in fruits) {
  print(i)
}
```
```{r}
fruits <- list("apple","banana","cherry")
for (i in fruits) {
  if(i== "banana"){
    break
  }
  print(i)
}

```
```{r}
fruits <- list("apple","banana","cherry")
for (i in fruits) {
  if(i== "banana"){
    next
  }
  print(i)
}
```
```{r}
dice <- 1:6
for (x in dice) {
  if(x==6){
    print(paste("The dice number is", x, " Six"))
  }else{
    print(paste("The dice number is", x, "Not Six"))
  }
}

```

### 25. Nested For Loops

```{r}
adj <- list(1:5)
name <- list("Delwar","Rakib","Akbar","Karim","Rahim")
for (i in adj) {
  for(j in name){
    print(paste(i,j))
  }
}
```

### 26. R Functions

```{r}
my_function <- function(){
  print("Hi")
}
my_function()

```
```{r}
my_function <- function(fname){
  print(fname)
}
my_function("Delwar")
my_function("Peter")
my_function("Shuvo")
```
```{r}
my_function <- function(fname,age){
  print(paste(fname,age))
}
my_function("Delwar",29)
my_function("Peter",27)
my_function("Shuvo",26)
```
### 27. Nested Functions 
There are two ways to create a nested function
```{r}
#call a function within another function

nested_function <- function(x,y){
  a <- x+y
  return(a)
}
nested_function(3,4)
nested_function(nested_function(3,4),nested_function(3,4))
```
```{r}
#write a function within a function 
out <- function(x){
  inner <- function(y){
    a<- x+y
    return(a)
  }
  return(inner)
}
result <- out(5)
result(3)

```

### 28.Function Recursion

```{r}
tri_function <- function(k){
  if(k>0){
    res <- k+tri_function(k-1)
    print(res)
  }else{
    res = 0
    return(res)
    
  }
}
tri_function(6)
```
### 29. Global Variable
Can be used by everyone both inside and outside
```{r}
txt <- "Awesome"
my_function <- function(){
  paste("R is", txt)
}
my_function()
```
```{r}
txt <- "Awesome"
my_function <- function(){
  txt = "Delwar"
  paste("R is", txt)
}
my_function()
txt
```
```{r}
txt <- "Awesome"
my_function <- function(){
  txt <<- "Delwar"  #global assignment operator
  paste("R is", txt)
}
my_function()
txt
```

## R Data Structure
### 1. R Vectors

```{r}
fruits <- c("banana","apple","orange")
fruits
```
```{r}
x = c(1,2,3,4,5)
x
```
```{r}
x = c(1:5)
x
z = 4:10
z
```
```{r}
# Vector with numerical decimals in a sequence
numbers1 <- 1.5:6.5
numbers1

# Vector with numerical decimals in a sequence where the last element is not used
numbers2 <- 1.5:6.3
numbers2

# Vector of logical values
log_values <- c(TRUE, FALSE, TRUE, FALSE)
log_values
```
```{r}
#vector length
f = c(20:30)
length(f)

#Sort a vector
x = c(4,7,4,12,5,8,3)
sort(x)

#access a vector
x[2]

#access multiple elements
x[c(2,6)]
```
```{r}
#change an item
x = c(4,7,4,12,5,8,3)
x[1] = 1
x

#repeat vector
x = rep(c(1,2,3), each=3)
x

x = rep(c(1,2,3), times=3)
x

x = rep(c(1,2,3), times = c(2,3,4))
x
```
```{r}
#generating sequence vector
num = 1:10
num

num = seq(from = 0, to= 10, by = 2)
num
```
### 2. R List 
 
```{r} 
i = list("apple","banana","grape")
i

#access list
i = list("apple","banana","grape")
i[2]

#change the value
i = list("apple","banana","grape")
i[2]= "palm"
i
```
```{r}
i = list("apple","banana","grape")
length(i)

#check items if exists
i = list("apple","banana","grape")
"banana" %in% i
"cake" %in% i
```
```{r}
#add list items
i = list("apple","banana","grape")
append(i,"orange")

i = list("apple","banana","grape")
append(i,"orange", after = 1)

```
```{r}
#remove list items
i = list("apple","banana","grape")
j = i[-1]
j

#range of index
l = list('a','b','r','t')
l
(l)[2:4]

#join two list
l3 = c(i,l)
l3
length(l3)
```
### 3. R Matrics

```{r}

```

### 4. R Arrays 

```{r}

```

### 5. R Data Frames 

```{r}

```
### 6. Factors 

```{r}

```
## R Graphics 

### 1. R Plot  

```{r}
plot(1,6)
```
```{r}
plot(c(1,6),c(3,8))
```
```{r}
plot(c(12,3,4,6),c(3,4,5,8))
```
```{r}
plot(1:10)
```
```{r}
plot(c(1,2,3,4,6),c(3,4,5,6,8), type = "l")
```
```{r}
plot(c(1,6),c(3,8), main = "Test", xlab = "x Test", ylab = "y Test")
```

```{r}
plot(1:10,col= "red",cex=2,pch=9, main = "Test", xlab = "x Test", ylab = "y Test")
```
### 2. R Line
```{r}
plot(1:10, type = "l", col ="blue", lwd = 3, lty = 5)
```

```{r}
line1 = seq(from = 1, to = 10, by=2)
line2 = seq(from =2, to= 10, by =2)
plot(line1, type = "l", col ="blue")
lines(line2,type = "l",  col = "red")
```

### 3. R Scatterplot
```{r}
x <- c(4,5,7,8,4,4,6,7,4,6,7,9)
y <- c(18,16,14,12,15,16,17,19,10,10,16,14)
plot(x,y)
```
```{r}
x <- c(4,5,7,8,4,4,6,7,4,6,7,9)
y <- c(18,16,14,12,15,16,17,19,10,10,16,14)
plot(x,y, main = "Observations of Cars", xlab = "Car Age", ylab = "Car Speed")
```
```{r}
x1 <- c(4,5,7,8,4,4,6,7,4,6,7,9)
y1 <- c(18,16,14,12,15,16,17,19,10,10,16,14)

x2 <- c(1,2,3,5,1,4,6,5,4,6,7,9)
y2 <- c(18,16,14,12,15,16,17,19,10,10,16,14)
plot(x1,y1, main = "Observations of Cars", xlab = "Car Age", ylab = "Car Speed", col = "red", cex= 2)
points(x2,y2, col = "blue", cex= 2)
```

### 4. R Pie Charts

```{r}
x <- c(10,20,30, 40)
pie(x)
```


```{r}
x <- c(10,20,30, 40)
pie(x, init.angle = 90)
```

```{r}
x <- c(10,20,30, 40)
label = c("Apple","Banana", "Orange","Grape")
color = c("green","yellow","orange","red")

pie(x, labels = label, main = "Fruits", col = color)

legend("bottomright", label, fill = color)
```

### 5. R Bars

```{r}
x <- c("A","B","C","D")
y <- c(2,4,6,8)
barplot(y, names.arg = x, col = "red", density = 10, width = rep(y))
```

```{r}
x <- c("A","B","C","D")
y <- c(2,4,6,8)
barplot(y, names.arg = x, col = "red", density = 10, horiz = TRUE)
```